Chemistry of fossil Pinus pollen and the effect of acetolysis

### 2021-05-28 manuscript.Rmd
rewrote introduction

In recent years, the study of pollen-grain chemistry has shown promising results for biological and palaeoecological research.
Fourier Transform Infrared (FTIR) spectroscopy has become a popular method  for characterising the chemical variation within pollen and spores.
A number of FTIR studies have identified pollen chemical differences on genus level, which may extent to species level differences [@zimmermannChemicalCharacterizationIdentification2018; @jardineChemotaxonomyDomesticatedGrasses2019; @muthreichChemicalVariationsQuercus2020; @zimmermannInfraredSpectroscopyPollen2014; @depciuchIdentificationBirchPollen2018; @diehnDiscriminationGrassPollen2020]. 
This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution @dellannaPollenDiscriminationClassification2009; @zimmermannAnalysisAllergenicPollen2016.
Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods [ @mondolApplicationHighThroughputScreening2019; @zimmermannAnalysisAllergenicPollen2016; @guedesPollenRamanSpectra2014].
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen [@bagciogluMultiscaleVibrationalSpectroscopic2015; @kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020].
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes.

So far, most FTIR studies focused on identification of pollen have used modern material and bulk measurements, while some used FTIR microscopes.
These FTIR microscopes required manual targeting of small areas of the sample slides to record spectra of single or groups of pollen grains.
FTIR as a method has advanced on a technical level over the past 5 to 10 years.
The development and capabilities of microscopes with connected FTIR spectrometers now allow easier and faster capture of spectra of pollen grains.
With these newer machines a larger area is irradiated with FTIR and spectra are recorded at higher resolution over the entire irradiated area.
This allows capture of an image and matrix of spectra over the same sample area, which enables e.g. the extraction of spectra from several single grains from the same area on the sample without the need to relocate the detector of the microscope or move the slide.

Palaeoecologists use a combination of several chemical extractions to isolate and concentrate pollen from the sediment matrix.
Most common is acetolysis in addition to HF treatment [@erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989].
These strong acid treatments aim to remove other organic particulates and silicates, while preserving morphological integrity of pollen grains [REF].
A limited number of studies have used acetolysis and other chemical treatments in combination with FTIR on pollen to simulate fossilization [@julierChemotaxonomyToolInterpreting2016; @jardinePollenSporesBiological2016; @woutersenNovelApproachStudy2018] or study the effects of acid extractions [@jardineImpactOxidationSpore2015; @jardineSporopolleninChemistryIts2021].
@jardineImpactOxidationSpore2015 found limited to no effect of short acetolysis (<10 min) on *Lycopodium* spore exine chemistry, while labile outer compounds and the intine/protoplast where removed.

An alternative to chemical extraction is density separation using e.g. sodium polytungstate (SPT), which is used to isolate pollen from minerogenic sediments [REFS] or as preparations for pollen isotope studies [REFS].
Density separation may be a an alternative to acid extractions for pollen studies concerned with chemical alteration of pollen grain chemistry.
SPT is an inorganic compound, which should not alter pollen chemistry in any way and is removed with several washes after extraction.
Contamination of the sample is easily identifiable by characteristic wavebands visible in the spectra.

@@

acetolysis alters sporopollenin [@dominguezIsolationIntactPollen1998; @schulzeosthoffPhenolsIntegratedCompounds1987] affecting solubility

>>>>

In recent years, the study of pollen-grain chemistry has shown promising results for biological and palaeoecological research.
Fourier Transform Infrared (FTIR) spectroscopy has become a popular method  for characterising the chemical variation within pollen and spores.
A number of FTIR studies have identified pollen chemical differences on genus level, which may extent to species level differences [@zimmermannChemicalCharacterizationIdentification2018; @jardineChemotaxonomyDomesticatedGrasses2019; @muthreichChemicalVariationsQuercus2020; @zimmermannInfraredSpectroscopyPollen2014; @depciuchIdentificationBirchPollen2018; @diehnDiscriminationGrassPollen2020]. 
This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution @dellannaPollenDiscriminationClassification2009; @zimmermannAnalysisAllergenicPollen2016.
<!--MOVE TO DISCUSSION: Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods [ @mondolApplicationHighThroughputScreening2019; @zimmermannAnalysisAllergenicPollen2016; @guedesPollenRamanSpectra2014].
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen [@bagciogluMultiscaleVibrationalSpectroscopic2015; @kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020].
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes. -->

So far, most FTIR studies focused on identification of pollen have used modern material and bulk measurements, while some have used FTIR microscopes.
These FTIR microscopes required manual targeting of small areas of the sample slides to record spectra of single grains or groups of pollen grains.
FTIR as a method has advanced on a technical level over the past 5 to 10 years.
The development and capabilities of microscopes with connected FTIR spectrometers now allow easier and faster capture of spectra of pollen grains.
With these newer machines a larger area is irradiated and spectra are recorded at higher resolution over the entire irradiated area.
This allows capture of an image and matrix of spectra over the same sample area, which enables e.g. the extraction of spectra from several single grains from the same area on the sample without the need to relocate the detector of the microscope or move the slide.

Palaeoecologists use a combination of several chemical extractions to isolate and concentrate pollen from the sediment matrix.
Most common is acetolysis in addition to HF treatment and HCl [@erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989].
These strong acid treatments aim to remove other organic particulates, calcium carbonates and silicates, while preserving morphological integrity of pollen grains [REF].
A limited number of studies have used acetolysis and other chemical treatments in combination with FTIR on pollen to simulate fossilization [@julierChemotaxonomyToolInterpreting2016; @jardinePollenSporesBiological2016; @woutersenNovelApproachStudy2018] or study the effects of acid extractions [@dominguezIsolationIntactPollen1998; @dominguezPollenSporopolleninDegradation1999; @loaderPreparationPollenStable2000;@schulzeosthoffPhenolsIntegratedCompounds1987;@jardineImpactOxidationSpore2015; @jardineSporopolleninChemistryIts2021].
These results suggest that acetolysis may alter the chemical composition of sporopollenin.

An alternative to chemical extraction is density separation using e.g. sodium polytungstate (SPT), which is used to isolate pollen from minerogenic sediments [REFS] or as preparations for pollen isotope studies [REFS].
Density separation may be a an alternative to acid extractions for pollen studies concerned with chemical alteration of sporomorph chemistry.
SPT is used to adjust the density of sample suspension and facilitace separation of less dense particulates (e.g. pollen and other microfossils) over mineral particles.
SPT is removed from the sample with several washes after extraction and any contamination of the sample is easily identifiable by characteristic wavebands in the spectra.

@@
